---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The `phscs` Package

<!-- badges: start -->
[![R-CMD-check](https://github.com/yng-me/phscs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yng-me/phscs/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/phscs)](https://CRAN.R-project.org/package=phscs)
<!-- badges: end -->

`phscs` provides a single, consistent interface to seven [Philippine Statistics Authority](https://psa.gov.ph/classification) (PSA) classification systems. All data is **bundled with the package** — no internet connection or API token required.

| Function | Classification |
|---|---|
| `get_psgc()` | Philippine Standard Geographic Code |
| `get_psic()` | Philippine Standard Industrial Classification |
| `get_psoc()` | Philippine Standard Occupational Classification |
| `get_psced()` | Philippine Standard Classification of Education |
| `get_pcoicop()` | Phil. Classification of Individual Consumption According to Purpose |
| `get_pcpc()` | Philippine Central Product Classification |
| `get_psccs()` | Philippine Standard Commodity Classification System |

## Installation

Install from CRAN:

``` r
install.packages("phscs")
```

Or install the development version from GitHub:

``` r
# install.packages("pak")
pak::pak("yng-me/phscs")
```

## Usage

```{r usage, message=FALSE}
library(phscs)

# Geographic data — re-exported from the psgc package
get_psgc(geographic_level = "region") |> head(3)

# Industrial classification (default: sub-classes)
get_psic() |> head(3)

# Occupational classification (default: unit groups)
get_psoc() |> head(3)
```

All classification functions share the same interface:

```r
# Choose a level of detail
get_psic(level = "sections")
get_psoc(level = "major")
get_pcoicop(level = "divisions")

# Include the full text description
get_psic(level = "sections", cols = "description")

# Switch to an older edition (where available)
get_pcoicop(version = "2009")
```

For a full walkthrough of every classification and more examples, see:

``` r
vignette("phscs")
```

